---
title: "scd_analysis"
author: "mateusz psujek"
date: "2023-07-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman)

pacman::p_load(tidyverse, rstanarm, brms, bayesplot, ggdist, ggthemes, patchwork, grid, ggsci, wesanderson)

theme_set(theme_tidybayes())
```
# to_do list

#### Defining the model
Model: multilevel varying-intercepts Bayesian binomial regression #to_do what type of binomial

Mathematical definition
$$
\begin{aligned}
NonO_i &\sim Binomial(N_i, p_i) \\
logit(p_i) &= \overline{\alpha} +\alpha_{\text{condition}[i]} + z_{\text{subject}[i]} \sigma_{subject} \\
\overline{\alpha} &\sim Normal(0, 1.5) \\
\alpha_{\text{condition}} &\sim Normal(0,0.5) \\
z_j &\sim Normal(0, 1) \;\;\; , \text{for } j = 1, \dots, 16 \\
\sigma_{subject} &\sim Expontential(1)
\end{aligned}
$$

### Loading the preprocessed data 
Pre-processing script can be found in the same directory in preprocessing.R file
```{r}
df_backgroud <- read_csv("background_data.csv") %>% 
  mutate(
    #reordering the conditions - useful later when plotting
    condition = factor(condition, levels = c("group", "individual"))
  )

df_ratio_bias <- read_csv("ratio_bias_data.csv")
df_invest_task <- read_csv("invest_task_data.csv")
df_belief_bias <- read_csv("belief_bias_data.csv")
```
### Joining the background data with each of the individual performance files
```{r}
join_background <- function(df){
    
    df <- df %>% 
      nest(.by = c(id, condition)) %>% 
      full_join(df_backgroud, by = c("id", "condition")) %>% 
      unnest(data)
  }
df_ratio_bias <- df_ratio_bias %>% join_background()
df_belief_bias <- df_belief_bias %>% join_background()
df_invest_task <- df_invest_task %>% join_background()
```

### Belief bias
```{r}
m_belief_bias_base <- 
  brm(data = df_belief_bias, 
      family = binomial,
      bf(nonoptimal | trials(1) ~ a + b,
         a ~ 1 + (1 | id),
         b ~ 0 + condition,
         nl = TRUE),
      prior = c(prior(normal(0, 0.5), nlpar = b),
                prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
                prior(exponential(1), class = sd, group = id, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/m_belief_bias_base")
```
#### checking convergance
```{r fig.width = 12}
print(m_belief_bias_base)
mcmc_trace(m_belief_bias_base)
```

### Investment task 

```{r}
m_invest_task_base <- 
  brm(data = df_invest_task, 
      family = binomial,
      bf(nonoptimal | trials(1) ~ a + b,
         a ~ 1 + (1 | id),
         b ~ 0 + condition,
         nl = TRUE),
      prior = c(prior(normal(0, 0.5), nlpar = b),
                prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
                prior(exponential(1), class = sd, group = id, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/m_invest_task_base")


```
```{r fig.width=12}
mcmc_trace(m_invest_task_base)
print(m_invest_task_base)
```
### Ratio bias

#### Some data wrangling
##### working out the interaction between number of red marbles in large tray (high(9,8) or low(7,6,5)) and the frame
```{r}
df_ratio_bias %>% glimpse

df_ratio_bias<- df_ratio_bias %>% 
  rowwise() %>% 
  mutate(
    current_frame = if_else(current_frame == "win_frame", "WF", "LF"),
    frame_magnitude = str_flatten(c(current_frame, "/", n_red_magnitude)) %>% as_factor,
    .after = nonoptimal
  ) %>% 
  filter(!n_red_large == 10)

df_ratio_bias %>% glimpse

distinct(df_ratio_bias, frame_magnitude)
```

#### Fitting the model
```{r}
# m_ratio_bias_base <- 
#   brm(data = df_ratio_bias, 
#       family = binomial,
#       bf(nonoptimal | trials(1) ~ a + b,
#          a ~ 1 + (1 | id),
#          b ~ 0 + condition + frame_magnitude,
#          nl = TRUE),
#       prior = c(prior(normal(0, 0.5), nlpar = b),
#                 prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
#                 prior(exponential(1), class = sd, group = id, nlpar = a)),
#       iter = 2000, warmup = 1000, chains = 4, cores = 4,
#       seed = 13,
#       file = "fits/m_ratio_bias_base")

m_ratio_bias_interaction <- 
  brm(data = df_ratio_bias, 
      family = binomial,
      bf(nonoptimal | trials(1) ~ a + b,
         a ~ 1 + (1 | id),
         b ~ 0 + condition + frame_magnitude,
         nl = TRUE),
      prior = c(prior(normal(0, 0.5), nlpar = b),
                prior(normal(0, 1.5), class = b, coef = Intercept, nlpar = a),
                prior(exponential(1), class = sd, group = id, nlpar = a)),
      iter = 2000, warmup = 1000, chains = 4, cores = 4,
      seed = 13,
      file = "fits/m_ratio_bias_interaction")
```



##### Convergance checks
```{r fig.width = 12}
mcmc_trace(m_ratio_bias_interaction)
print(m_ratio_bias_interaction)
```
```{r}
m_ratio_bias_interaction %>% 
  mcmc_plot(variable = c("^b_", "^sd_"), regex = T) +
  theme(axis.text.y = element_text(hjust = 0))
```
### Plotting

##### the contrasts
```{r}

# replace here if you change models!
models <- list(m_belief_bias_base, m_invest_task_base, m_ratio_bias_interaction)
names(models) <- c("belief_bias", "invest_task", "ratio_bias")

posts <- map(models, as_draws_df)


plots_contrast <- map(
  posts,
  .f = function(post){
      df_contrast <- post %>% 
        mutate(
        contrast = b_b_conditionindividual - b_b_conditiongroup,
        contrast_p = inv_logit_scaled(b_b_conditionindividual) - inv_logit_scaled(b_b_conditiongroup)
         ) %>% 
      select(contains("contrast")) %>% 
      pivot_longer(everything())
    
      
      plot_logodds <- 
        df_contrast %>% 
        filter(name == "contrast") %>% 
        ggplot(aes(x = value, y = '')) +
        stat_halfeye(fill = "darkgrey", alpha = 0.7) +
        geom_vline(xintercept = 0, linetype = 'dashed', color = 'darkred') +
        labs(
          y = NULL,
          x = "Posterior log-odds"
          )
      
      plot_prob <- 
        df_contrast %>% 
        filter(name == "contrast_p") %>% 
        ggplot(aes(x = value, y = '')) +
        stat_halfeye(fill = "darkgrey", alpha = 0.7) +
        geom_vline(xintercept = 0, linetype = 'dashed', color = 'darkred') +
        labs(
          y = NULL,
          x = "Posterior on probablity scale"
          )
      output <- list(plot_logodds, plot_prob)
      names(output) <- c("logodds", "prob")
      return(output)
  }
)

print(plots_contrast)
```
#### plotting the posterior predictions
```{r}
new_data <- distinct(df_belief_bias, condition)

belief_bias_average <- 
  fitted(m_belief_bias_base,
         newdata = new_data,
         re_formula = NA,
         probs = c(.1, .9)
         ) %>% 
  data.frame() %>% 
  bind_cols(new_data) %>% 
  mutate(type = "average")




belief_bias_marginal <- 
  fitted(m_belief_bias_base,
         newdata = new_data,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         probs = c(.1, .9)
         ) %>% 
  data.frame() %>% 
  bind_cols(new_data) %>% 
  mutate(type = "marginal")

#average
plot_belief_bias_average <- belief_bias_average %>% 
  mutate(condition = factor(condition, levels = c("individual", "group"))) %>% 
  ggplot(aes(x = Estimate, y = condition, group = 1)) +
  geom_pointinterval(aes(xmin = Q10, xmax = Q90)) +
  geom_line() +
  coord_cartesian(xlim = 0:1) +
  labs(
    y = NULL,
    x = "Probability of non-optimal response"
    )
plot_belief_bias_average

#marginal
plot_belief_bias_marginal <- belief_bias_marginal %>% 
  mutate(condition = factor(condition, levels = c("individual", "group"))) %>% 
  ggplot(aes(x = Estimate, y = condition, group = 1)) +
  geom_pointinterval(aes(xmin = Q10, xmax = Q90)) +
  geom_line() +
  coord_cartesian(xlim = 0:1)+
    labs(
      y = NULL,
      x = "Probability of non-optimal response"
      )
plot_belief_bias_marginal

#both average and marginal wrapped next to each other
bind_rows(belief_bias_average, belief_bias_marginal) %>% 
  ggplot(aes(x = Estimate, y = condition, group = 1))+
  geom_pointinterval(aes(xmin = Q10, xmax = Q90)) +
  geom_line()+
  coord_cartesian(xlim = 0:1) +
  facet_grid(rows = vars(type))
```


```{r}
new_data <- distinct(df_invest_task, condition)

plot_invest_task_marginal <- 
  fitted(m_invest_task_base,
         newdata = new_data,
         allow_new_levels = T,
         sample_new_levels = "gaussian",
         probs = c(.1, .9)
         ) %>% 
  data.frame() %>% 
  bind_cols(new_data) %>% 
  mutate(condition = factor(condition, levels = c("individual", "group"))) %>% 
  ggplot(aes(x = Estimate, y = condition, group = 1)) +
  geom_pointinterval(aes(xmin = Q10, xmax = Q90)) +
  geom_line() +
  coord_cartesian(xlim = 0:1)+
    labs(
      y = NULL,
      x = "Probability of non-optimal response"
      )
plot_invest_task_marginal
```
#### ratio bias predictive plot
```{r}
new_data <- expand_grid(
  distinct(df_ratio_bias, condition),
  distinct(df_ratio_bias, frame_magnitude)
)

d <- fitted(
  m_ratio_bias_interaction,
  newdata = new_data,
  allow_new_levels = T,
  sample_new_levels = "gaussian",
  probs = c(.1, .9) 
  ) %>% 
  data.frame() %>% 
  bind_cols(new_data) %>% 
  arrange(desc(Estimate)) %>% 
  mutate(frame_magnitude = factor(frame_magnitude, unique(frame_magnitude)))

plot_ratio_bias_marginal <- d %>% 
  ggplot(aes(x = Estimate, y = frame_magnitude, group = condition, color = condition)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(xmin = Q10, xmax = Q90), alpha = 0.8, width = 0.2) +
  coord_cartesian(xlim = 0:1)+
  labs(
    y = NULL,
    x = "Probability of non-optimal response"
  ) +
  scale_color_lancet()
plot_ratio_bias_marginal
  
```





### Making the final figure for the paper
```{r fig.width=10}

l1 <- wrap_elements(panel = textGrob('Belief bias', rot=90, vjust = 3))
l2 <- wrap_elements(panel = textGrob('Ratio bias', rot=90, vjust = 3))
l3 <- wrap_elements(panel = textGrob('Investment task', rot=90, vjust = 3))

(l2 | plots_contrast[["ratio_bias"]][["logodds"]] + labs(x = NULL) | plot_ratio_bias_marginal + labs(x = NULL)) /
(l1 | plots_contrast[["belief_bias"]][["logodds"]] + labs(x = NULL) | plot_belief_bias_marginal + labs(x = NULL)) /
(l3 | plots_contrast[["invest_task"]][["logodds"]] | plot_invest_task_marginal)


(l2 | plots_contrast[["ratio_bias"]][["logodds"]] + labs(x = NULL) )/
(l1 | plots_contrast[["belief_bias"]][["logodds"]] + labs(x = NULL) ) /
(l3 | plots_contrast[["invest_task"]][["logodds"]])

blank_label_contrast <- wrap_elements(panel = textGrob('Posterior log-odds', vjust = -4))
blank_label_marginal <- wrap_elements(panel = textGrob('Probability scale', vjust = -4))


(plots_contrast[["ratio_bias"]][["logodds"]] + 
    labs(
      title = "Ratio bias",
      x = NULL
      ) | 
 plots_contrast[["belief_bias"]][["logodds"]] + 
    labs(
      title = "Belief bias",
      x = NULL
      ) | 
 plots_contrast[["invest_task"]][["logodds"]] + 
    labs(
      title = "Investment task",
      x = NULL
      ) 
  ) / 
  blank_label_contrast

ggsave("plots/fig_contrast(ind-group).png")


(plot_ratio_bias_marginal + 
    labs(
      title = "Ratio bias",
      x = NULL
      ) +
    theme(legend.position = "top") | 
 plot_belief_bias_marginal + 
    labs(
      title = "Belief bias",
      x = NULL
      ) | 
 plot_invest_task_marginal + 
    labs(
      title = "Investment task",
      x = NULL
      ) 
  ) / 
  blank_label_marginal

ggsave("plots/fig_marginal.png")

```

